---
title: "Week 5 - GGPLOT Revisited"
author: "Tim Kelleman"
date: "2024-02-22"

output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    fig_width: 8
    fig_caption: yes
    number_sections: yes
    toc_collapsed: yes
    code_folding: hide
    code_download: yes
    smooth_scroll: true
    theme: readable
   
  pdf_document: 
    toc: yes
    toc_depth: 4
    fig_caption: yes
    number_sections: yes
    fig_width: 5
    fig_height: 4
---

```{=html}
<style type="text/css">

div#TOC li {
    list-style:none;
    background-color:lightgray;
    background-image:none;
    background-repeat:none;
    background-position:0;
    font-family: Arial, Helvetica, sans-serif;
    color: #780c0c;
}

/* mouse over link */
div#TOC a:hover {
  color: red;
}

/* unvisited link */
div#TOC a:link {
  color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 {
    font-size: 24px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 {
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { 
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}

/* unvisited link */
a:link {
  color: green;
}

/* visited link */
a:visited {
  color: green;
}

/* mouse over link */
a:hover {
  color: red;
}

/* selected link */
a:active {
  color: yellow;
}

</style>
```
---
```{r setup, include=FALSE}
# code chunk specifies whether the R code, warnings, and output 
# will be included in the output files.
options(repos = list(CRAN="http://cran.rstudio.com/"))
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")             # Install png package
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")           # Install RCurl package
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
### ggplot and extensions
if (!require("ggplot2")) {
    install.packages("ggplot2")              
    library("ggplot2")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("ggridges")) {
    install.packages("ggridges")              
    library("ggridges")
}
if (!require("graphics")) {
    install.packages("graphics")              
    library("graphics")
}
if (!require("dplyr")) {
    install.packages("dplyr")              
    library("dplyr")
}
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE,
                      comment = NA)
```

```{r eval=TRUE}
#Import data
Countries<- read.csv("https://tkelleman.github.io/tkweb/Week5/countries_total.csv")
Income<-read.csv("https://tkelleman.github.io/tkweb/Week5/income_per_person.csv")
LifeExp<-read.csv("https://tkelleman.github.io/tkweb/Week5/life_expectancy_years.csv")
Population<-read.csv("https://tkelleman.github.io/tkweb/Week5/population_total.csv")

## A - Reshape data set: Income Per Person to make a longitudinal data such that the resulting data set has three columns: country, year, and income.
IncomeEdit <- Income %>%
  gather(key = "Year", value = "Income", - geo, na.rm=TRUE)
names(IncomeEdit)[1] <- "Country"

## B - Do the same for Life Expectancy in Years so that the resulting data set has three columns: country, year, and life expectancy.
LifeExpEdit <- LifeExp %>%
  gather(key = "Year", value = "LifeExpectancy", - geo, na.rm=TRUE)
names(LifeExpEdit)[1] <- "Country"

## C -  Merge/join the above two longitudinal data sets to make a new data set, under name LifeExpIncom that has variables: country, year, lifeExp, and income.
LifeExpIncom <- merge(IncomeEdit, LifeExpEdit, by = c("Country", "Year"))

## D - Merge LifeExpIncom with country region so that the final data set has information about income, life expectancy, and country region.
LifeIncomCount <-merge(LifeExpIncom, Countries, by.x = "Country", by.y = "name", all.x = TRUE)

## E -  Merge the previous resulting data set with population size so that the final data set has information about income, life expectancy, population size, and country region.
PopulationEdit <- Population %>%
  gather(key = "Year", value = "Population", - geo, na.rm=TRUE)
names(PopulationEdit)[1] <- "Country"

FinalData <- merge(LifeIncomCount, PopulationEdit, by = c("Country", "Year"))
FinalData$Year<-gsub("X","", as.character(FinalData$Year))
#write.csv(FinalData, "FinalData.csv")

## 3 Create a subset of the above resulting longitudinal data set that contains only the data of the year 2000/font> - name it 2000data.
Y2000data<-filter(FinalData, Year==2000)

```

---
```{r eval=TRUE}
cols1 = c("#332288","#117733","#44AA99","#88CCEE","#DDCC77","#CC6677")
Y2000Plot<-ggplot(Y2000data, aes(x=Income, 
                      y=LifeExpectancy, 
                      size = Population, 
                      color = region)) +
  geom_point(aes(size=Population, ids = Country),
             show.legend = TRUE,
             alpha = 0.8)+
    scale_color_manual(values = cols1)+
    labs(title = 'Plot of Life Expectancy by Income in 2000',
       x = 'Income',
       y = 'Life Expectancy')+
  transition_states(Year) +
  ease_aes('linear')
Y2000Plot
```
Figure 1: Plot of Life Expectancy by Income in 2000 \n This plot shows the relationship between life expectancy and income in the year 2000 in the merged data set. Population of each country is represented by the diameter of the points plotted and the color of the point represents the region of the country. As shown in this figure, countries in the regions of Americas and Europe have a higher life expectancy than most countries in the Africa region, with a few outlier data points.
---
```{r eval=TRUE}
### Full Data Set
cols1 = c("#332288","#117733","#44AA99","#88CCEE","#DDCC77","#CC6677")
AllDataPlot<-ggplot(FinalData, aes(x=Income, 
                      y=LifeExpectancy, 
                      size = Population, 
                      color = region)) +
  geom_point(aes(size=Population, ids = Country),
             show.legend = TRUE,
             alpha = 0.8)+
    scale_color_manual(values = cols1)+
    labs(title = 'Plot of Life Expectency by Income at Year: {closest_state}',
       x = 'Income',
       y = 'Life Expectancy')+
  transition_states(Year) +
  ease_aes('linear')

animate(AllDataPlot, renderer = gifski_renderer(), rewind = FALSE, nframes = 300, fps = 10)
```
Figure 2: Plot of Life Expectancy by Income for Years 1800-2018 \n This plot shows the relationship between life expectancy and income all countries in the merged data set. Population of each country is represented by the diameter of the points plotted and the color of the point represents the region of the country. As shown in the plot, all represented countries increase their life expectancy between the given years but not all countries show an increase in income. 




